---
title: "Shiny_Lec6"
author: "Jingyang(Judy) Zhang"
date: "5/20/2020"
output: html_document
---

## Lession 6 Use Reactive Expressions ##

1. Reactive expressions let you control which parts of your app update when, which prevents unnecessary computation that can slow down your app. 


### A New App: stockVis

1. The stockVis app looks up stock prices by ticker symbol and displays the results as a lne chart. The app lets you:

a. Select a stock to examine

b. Pick a range of dates to review

c. Choose whether to plot stock prices or the log of the stock prices on the y axis

d. Decide whether or not to correct prices for inflation

e. By default, stock Vis displays the SPY ticker (an index of the entire S$P 500). To look up a different stock, type in a stock symbol that Yahoo finance will recognize. 

2. stockVis relies heavily on two functions from the `quantmod` package:

a. It uses `getSymbols` to download financial data straight into R from websites like Yahoo finance and the Federal Reserve Bank of St. Louis. 

b. It uses `chartSeries` to display prices in an attractive chart. 

3. stockVis also relies on an R script named `helpers.R`, which contains a function that adjusts stock prices for inflation. 


### Check Boxes and Date Ranges ###

1. The stockVis app uses a few new widgets:

a. A date range selector, created with `dateRangeInput`

b. A couple of check boxes made `checkboxInput`. Check box widgets are very simple. They return a `TRUE` when the checkbox is checked, and a `FALSE` when the checkbox is not checked. 


### Streamline Computation ###

1. Plot y axis on the log scale Checkbox: the value of `input$log` will change, which will cause the entire expression in `renderPlot` to rerun:

```{r eval = FALSE}

output$plot <- renderPlot({
  data <- getSymbols(input$symb, src = "yahoo",
                     from = input$dates[1],
                     to = input$dates[2],
                     auto.assign = FALSE)
  chartSeries(data, theme = chartTheme("white"),
              type = "line", log.scale = input$log, TA = NULL)
  
  
})
```


2. Each time `renderPlot` reruns:

a. It re-fetches the data from Yahoo finance with `getSymbols` 

b. It re-draws the cahrt with the correct axis

This is NOT good, because you do NOT need to re-fetch the data to re-draw the plot. In fact, Yahoo finance will cut you off if you re-fetch your data too often (because you begin to look like a bot). 

But more importantly, re-running `getSymbols` is unnecessary work, which can slow down your app and consume server bandwidth. 


### Reactive Expressions ###

1. You can limit what gets re-run during a reaction with reactive expressions. 

a. Reative Expression: an R expression that uses widget input and return a value. The reactive expression will update this value whenever the original widget changes. 

b. To create a reactive use the `reactive` function, which takes an R expression surrounded by braces (just like the `render*` functions)


__Example__: Reactive Expression 

```{r eval = FALSE}


# A reactive expression that uses the widgets of stockVis to fetch data from Yahoo.
dataInput <- reactive({
  getSymbols(input$symb, src = "yahoo",
             from = input$dates[1],
             to = input$dates[2],
             auto.assign = FALSE)
})

```

When you run the expression, it will run `getSymbols` and return the results, a data frame of price data. You can use the expression to access price data in `renderPlot` by calling `dataInput()`. 

```{r eval = FALSE}

output$plot <- renderPlot({
  chartSeries(dataInput(), theme = chartTheme("white"),
              type = "line", log.scale = input$log, TA = NULL)
  
})

```


2. Reactive expression are a bit smarter than regular R functions. They cache their values and know when their values have become outdated. 

a. The first time that you run a reactive expression, the expression will save its result in your computer's memory. 

b. The next time you call the reactive expression, it can return this saved result without doing any computation (which make your app faster). 

c. The areactive expression will only return the saved result if it knows that the result is up-to-date. If the reactive expression has learned that the result is obsolete (because a widget has changed), the expression will recalculate the result. It then returns the new result and saves a new copy. The reactive expression will use this new copy until it too becomes out of date. 


___Summary___: Reactive Expression

1. A reactive expression saves its result the first time you run it. 

2. The next time the reactive expression is called, it checks if the saved value has become out of date (i.e. Whether the widgets it depends on have changed).

3. If the value is out of date: the reactive object will recalculate and then save the new result.

4. If the value is up to date: the reactive expression will return the saved value without doing any computation. 


__Example__: stockVis Reactive Expression

```{r eval = FALSE}

server <- function(input, output) {
  dataInput <- reactive({
    getSymbols(input$symb, src = "yahoo",
               from = input$dates[1],
               to = input$dates[2],
               auto.assign = FALSE)
  })
  
  output$plot <- renderPlot({
    
    chartSeries(dataInput(), theme = chartTheme("white"),
                type = "line",
                log.scale = input$log, TA = NULL)
  })
}

```

When you click "Plot y axis on the log scale", `input$log` will change and `renderPlot` will re-execute. Now:

a. `renderPlot` will call `dataInput()`

b. `dataInput` will check that the `dates` and `symb` widgets have not changed

c. `dataInput` will return its saved data set of stock prices without re-fectching data from Yahoo

d. `renderPlot` will re-draw the chart with the correct axis



### Dependencies ###

1. Shiny keeps track of which reactive expressions and `output` object depends on, as well as which widget inputs. Shiny will automatically re-build an object if:

a. An `input` value in the objects's `render*` function changes OR

b. A reactive expression in the objects's `render*` function becomes obsolete

2. Reactive expressions are like links in a chain that connect `input` values to `output` objects. The objects in `output` will respond to changes made anywhere downstream in the chain. 

3. Only call a reactive expression from within a `reactive` or a `render*` function as these R functions are equipped to deal with reactive output, which can change without warning. 

In fact, Shiny will prevent you from calling reactive expressions outside of these functions. 


__Answer 1__

```{r eval = FALSE}
server <- function(input, output) {

  dataInput <- reactive({  
      getSymbols(input$symb, src = "yahoo",
          from = input$dates[1],
          to = input$dates[2],
          auto.assign = FALSE)
  })

  finalInput <- reactive({
    if (!input$adjust) return(dataInput())
    adjust(dataInput())
  })

  output$plot <- renderPlot({
    chartSeries(finalInput(), theme = chartTheme("white"),
        type = "line", log.scale = input$log, TA = NULL)
  })
}
```


The above code isolates each input in its own reactive expression or `render*` function. If an input changes, only out of date expressions will re-run. 

The flow:
1. A user clicks "Plot y axis on the log scale"

2. `renderPlot` re-runs

3. `renderPlot` calls `finalInput`

4 `finalInput` checks with `dataInput` and `input$adjust`. 

5. If neither has changed, `finalInput` returns its saved value. 

6. If either has changed, `finalInput` calculates a new value with the current inputs. It will pass the new value to `renderPlot` and store the new value for future queries. 


### Recap ###

1. YOu can make your app faster by modularizing your code with reactive expressions:

a. A reactive expression takes `input` values, or values from other reactive expressions, and returns a new value

b. Reactive expressions save their results and will only re-calculate if their input has changed. 

c. Create reactive expressions with `reactive({})`

d. Call reactive expressions with the name of the expression followed by parenthese `()`

e. Only call reactive expressions from within other reactive expressions or `render*` functions. 