---
title: "Shiny_Lec1"
author: "Jingyang(Judy) Zhang"
date: "5/18/2020"
output: html_document
---

## Lession 1 Welcome to Shiny ##

1. Shiny is an R package that makes it easy to build interactive web applications (apps) straight from R. 

2. The Shiny package contains 11 built-in examples that each demostrate how Shiny works. 

a. Each example is a self-contained Shiny app. 

b. The Hellow Shiny example plots a histogram of R's `faithful` dataset with a configurable number of bins. 

(1) User can change the number of bins with a slider bar, and the app will immediately respond to their input. 

(2) To run Hello Shiny:
```{r}
library(shiny)
runExample("01_hello")
```


### Structure of a Shiny App ###

1. Shiny apps are contained in a single script called `app.R`. 

a. The script `app.R` lives in a director (for example , `newdir/`) and the app can be run with `runApp("newDir")`. 

b. `app.R` has 3 components:

(1) a user interface (ui) object: controls the layout and appearance of your app. 

(2) a server function: contains the instructions that your computer needs to build your app. 

(3) a call to the `shinyApp` function: creates Shiny app objects from an explicit UI/server pair. 

__Note__: Prior to version 0.10.2, Shiny did NOT support single-file apps and the `ui` object and `server` function needed to be contained in separate scripts called `ui.R` and `server.R`, respectively. This functionality is still supported in Shiny. 

2. One nice feature about single-file apps is that you can copy and paste the entire app into the R console, which makes it easy to quickly share code for others to experiment with. 


__ui__: the `ui` object for the Hello Shiny example.

```{r}

# Define UI for app that draws a histogram ----

ui <- fluidPage(
  
  # App title ----
  titlePanel("Hello Shiny!"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel (
      
      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
    ),
    
    # Main panel for displaying outputs ----
    mainPanel (
      
      # Output: Histogram ----
      plotOutput(outputId = "distPlot")
    )
  )
)


```




__Server__: the `server` function for the Hello Shiny example.

```{r}
# Define server logic required to draw a histogram ----

server <- function(input, output){
  
  
  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call 
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #   re-executed when inputs (input$bins) change
  # 2. Its ouput type if a plot
  
  output$distPlot <- renderPlot({
    
    e <- faithful$waiting
    
    bin <- seq(min(x), max(x), length.out = input$bins + 1)
    
    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")
  })
  
  
}
```



3. Note that in your `app.R` file you will eed to start with loading the Shiny package and end with a cal to `shinyApp`:

```{r eval=FALSE}

library(shiny)

ui <- ...

server <- ...

shinyApp(ui = ui, server = server)
)

```



4. Your R session will be busy while the Hello Shiny app is active, so you will NOT be able to run any R commands. R is monitoring the app and executing the app's reactions. 

a. To get your R session back, hit escape or click the stop sign icon found in the upper right corner of the RStudio console panel. 

### Running an App ###

1. Every Shiny app has the same structure: an `app.R` file that contains `ui` and `server`. You can create a Shiny app by making a new directory and save an `app.R` file inside it. 

a. It is recommended that each app will live in its own unique directory. 

2. You can run a Shiny app by giving the name of its directory to the function `runApp`. 

```{r eval=FALSE}
library(shiny)

runApp("my_app")
```



3. By default, Shiny apps display in "normal" mode. 

a. Hellow Shiny and other built-in examples display in "showcase mode", a different mode that displays the `app.R` script alongside the app.

(1) Showcase Mode: display the app along with the script. 

b. If you would like your app to display in showcase mode, you can run `runApp("App-1", display.mode = "showcase")`.


### Relaunching Apps ###

1. To relaunch your Shiny app:

a. Run `runApp("App-1")` OR

b. Open the `app.R` script in your RStudio editor. RStudio will recognize the Shiny script and provide a Run App button (at the top of the editor). Either click this button to launch your app or use the keyboard shortcut: Command+Shift+Enter (Control+Shift+Enter on Windows)

(1) RStudio will launch the ap in a new window by default, but you can also choose to have the app launch in a dedicated viewer pane, or in your external web browser by clicking the icon next to Run App. 


### Recap ###
1. To create your own Shiny app:

a. Make a directory named `myapp/` for your app.

b. Save your `app.R` script inside that directory.

c. Launch the app with `runApp` or RStudio's keyboard shortcuts. 

d. Exit the Shiny app by clicking escape. 

### Go Further ###
1. You can create Shiny apps by copying and modifying exisitng Shiny apps. The Shiny gallery provides some good examples, or use the 11 pre-built Shiny examples lited below.

```{r}

runExample("01_hello")      # a histogram
runExample("02_text")       # tables and data frames
runExample("03_reactivity") # a reactive expression
runExample("04_mpg")        # global variables
runExample("05_sliders")    # slider bars
runExample("06_tabsets")    # tabbed panels
runExample("07_widgets")    # help text and submit buttons
runExample("08_html")       # Shiny app built from HTML
runExample("09_upload")     # file upload wizard
runExample("10_download")   # file download wizard
runExample("11_timer")      # an automated timer

```

