---
title: "L09 Themes"
subtitle: "Data Visualization (STAT 302)"
author: "YOUR NAME"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r global_options, include = FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE, dpi = 300)
```

## Overview

The goal of this lab is to play around with the theme options in `ggplot2` 

## Datasets

We'll be using the `cdc.txt` datasets.

<br>

```{r load-packages-data-seed, eval=FALSE}
# Load package(s)

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor"),
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")
  ))

# Set seed
set.seed(8221984)

# Selecting a random subset of size 100
cdc_small <- cdc %>% sample_n(100)
```

<br>

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Use the `cdc_small` dataset to explore several pre-set `ggthemes`. The code below constructs the familiar scatterplot of `weight` by `height` and stores it in `plot_01`. Display `plot_01` to observe the default theme. **Explore/apply, and display at least 7 other pre-set themes** from the `ggplot2` or `ggthemes` package. Don't worry about making adjustments to the figures under the new themes. Just get a sense of what the themes are doing to the original figure `plot_01`. 

Which theme or themes do you particularly like? Why? There should be at least 8 plots for this task. `temp1` is pictured below.

<br>

```{r base-plot-ex01}
# Building plot
plot_01 <- ggplot(data = cdc_small, aes(x = height, y = weight)) +
  geom_point(size = 3, aes(shape = genhlth, color = genhlth)) +
  scale_y_continuous(
    name = "Weight in Pounds",
    limits = c(100, 300),
    breaks = seq(100, 350, 25),
    trans = "log10",
    labels = scales::unit_format(unit = "lbs")
  ) +
  scale_x_continuous(
    name = "Height in Inches",
    limits = c(55, 80),
    breaks = seq(55, 80, 5),
    labels = scales::unit_format(unit = "in", accuracy = 1)
  ) +
  scale_shape_manual(
    name = "General\nHealth?",
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"),
    values = c(17, 19, 15, 9, 4)
  ) +
  scale_color_brewer(
    name = "General\nHealth?",
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"),
    palette = "Set1"
  ) +
  theme(
    legend.position = c(1, 0),
    legend.justification = c(1, 0)
  ) +
  labs(title = "CDC BRFSS: Weight by Height")
```

<br>

### Exercise 2

Using `plot_01` from  and the `theme()` function, attempt to construct the ugliest plot possible (example pictured below).

<br>



<br>

### Exercise 3

Using data from `cdc_small` create a few (at least two) graphics (maybe one scatterplot and one barplot). Style the plots so they follow a "Northwestern" theme. Check out the following webpages to help create the theme:

[Visual Identity](https://www.northwestern.edu/brand/visual-identity/index.html)

- [Web Colors](https://www.northwestern.edu/brand/visual-identity/color/digital-colors.html)
- [Secondary Palette](https://www.northwestern.edu/brand/visual-identity/color/secondary-palette.html)
- [Fonts & Typography](https://www.northwestern.edu/brand/visual-identity/fonts-typography/index.html) --- Need to use [substitute fonts](https://www.northwestern.edu/brand/visual-identity/fonts-typography/substitute-fonts.html)

<br>
