---
title: "L05 Toolbox"
subtitle: "Data Visualization (STAT 302)"
author: Jingyang Zhang
output:
  html_document:
    toc: true
    toc_float: true
---


## Set resolution
```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300)
```



## Overview

The goal of this lab is to explore more useful plots in `ggplot2`. Specifically we will be focusing on surface plots and geospatial plots (maps).

Challenges are not mandatory for students to complete. We highly recommend students attempt them though. *We would expect graduate students to attempt the challenges.*



## Datasets

We'll be using data from the `blue_jays.rda`, `cdc.txt`, and the `US_income.rda` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. 

```{r load-packages, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(viridis)
library(sf)
library(statebins)
library(tidycensus)
library(mapview)
options(tigris_use_cache = TRUE)

# Load datasets
load("data/blue_jays.rda")

load("data/US_income.rda")

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))
```


## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using `blue_jays.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- `geom_density_2d()` or `stat_density_2d()`
- Transparency is 1/3
- Horizontal lower limit is 57 and upper limit is 82
- Point size 1.5
- Line size is 0.4
- `binwidth` set to 0.004
- Minimal theme

```{r exer1}

ggplot(blue_jays, aes(Mass, Head)) +
  geom_point(alpha = 1/3, size = 1.5) +
  geom_density_2d(size = 0.4, color = "black", binwidth = 0.004) +
  xlim(57, 82) +
  labs(x = "Body mass (g)", y = "Head length (mm)") +
  theme_minimal()

```



<br>

### Exercise 2

Using `cdc` dataset, recreate the following graphics as precisely as possible.

##### Plot 1

*Hints:*

- `bins` set to 35
- Minimal theme

```{r exer2-p1}

ggplot(cdc, aes(height, weight)) +
  geom_hex(bins = 35) +
  labs(x = "Height (in)", y = "Weight (lbs)") +
  theme_minimal()

# use geom_hex(color="white") to add a white border
  
```


<br>

##### Plot 2

*Hints:*

- `polygon`
- Minimal theme

```{r exer2-p2}

ggplot(cdc, aes(height, weight)) +
  stat_density_2d(aes(fill=..level..), geom="polygon", show.legend = FALSE) +
  labs(x = "Height (in)", y = "Weight (lbs)") +
  facet_wrap(~gender) +
  theme_minimal()

# ..level.. is the same as after_stat(level) which will do things after the statistical transformation

```


<br>

### Exercise 3

Make a county map of a US state. Maybe use your home state (mine is below) or a favorite state. Please do not use the the state in example provided in the book

*Hints:*

- See section 3.7.1 in *ggplot2*
- Void theme

```{r get-map-data}

sd_counties <- map_data("county", "South Dakota") %>%
  select(lon = long, lat, group, id = subregion)

head(sd_counties)

```



```{r vis-sd}


ggplot(sd_counties, aes(lon, lat)) +
  geom_polygon(aes(group = group), fill = NA, color = "grey50") +
  coord_quickmap() +
  ggtitle("South Dakota") +
  theme_void()
  
  

```

<br>

### Exercise 4

Using `US_income` dataset, recreate the following graphics as precisely as possible.

<br>

```{r}
# Setting income levels
US_income <- mutate(
  US_income,
  income_bins = cut(
    ifelse(is.na(median_income), 25000, median_income),
    breaks = c(0, 40000, 50000, 60000, 70000, 80000),
    labels = c("< $40k", "$40k to $50k", "$50k to $60k", "$60k to $70k", "> $70k"),
    right = FALSE
  )
)

```

<br>


##### Plot 1

*Hints:*

- `geom_sf()` --- boundary color is `"grey80"` and size is 0.2
- `viridis` package (`discrete = TRUE` in `scale_*` function)
- Void theme

```{r get-state-geo}

# geometry in the US_income dataset contains all the geo information needed to draws states

ggplot(US_income) +
  geom_sf(aes(geometry = geometry, fill = income_bins), color = "grey80", size = 0.2) +
  scale_fill_viridis(discrete = TRUE, name = "Median\nIncome") +
  theme_void()

```


<br>

##### Plot 2

*Hints:*

- [Download developmental version](https://github.com/hrbrmstr/statebins) of `statebins` package
- `geom_statebins()`
- `viridis` package (`discrete = TRUE` in `scale_*` function)
- Statebins theme

Use shift+click to open the link

```{r}

ggplot(US_income, aes(state = name, fill = income_bins)) +
  geom_statebins() +
  scale_fill_viridis(
    discrete = TRUE,
    name = "Median\nIncome"
  ) +
  theme_statebins()

# discrete = TRUE: generates discrete color palette

# scale function is set to be scale_[argument you are working with, i.e. fill, color, etc.]+[pkg name]

# scale_fill_viridis can have the argument: option="magma" or sth else

# labs() function sets all the labels at once. 


```


<br>

## Challenge(s)

Using the `tidycensus` package and few others, try to create a map like below using [these directions](https://walkerke.github.io/tidycensus/articles/spatial-data.html). Try using a different geographical area and a different variable from the ACS.

```{r}


#census_api_key("faab61c2f4527d72cdcc4197c970bc831da4b50e", install=TRUE)

ny <- get_acs(geography = "tract", 
              variables = "B19013_001", 
              state = "NY", 
              county = "New York", 
              geometry = TRUE)

mapview(ny, zcol = "estimate", legend = TRUE)

```



<br>

