---
title: "L04 ggplot III"
author: "Data Visualization (STAT 302)"
output:
  html_document:
    toc: true
    toc_float: true
---

# Set resolution
```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300)
```

# Overview

The goal of this lab is to continue the process of unlocking the power of `ggplot2` through constructing and experimenting with a few basic plots.

# Datasets

We'll be using data from the `blue_jays.rda`, `tech_stocks.rda`, `corruption.rda`, and `cdc.txt` datasets which are already in the `/data` subdirectory in our **data_vis_labs** project. 

```{r load-packages, eval=TRUE, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)

library(ggrepel)


# Load datasets
load("data/blue_jays.rda")

load("data/tech_stocks.rda")

load("data/corruption.rda")

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))


```

# Exercises

Complete the following exercises.

<br>

### Exercise 1

Using `blue_jays.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Transparency is 0.6
- Point size 2
- Shift text horizontally by 0.5
- See pg 44 in *ggplot2*

<br>

```{r}

xrng=range(blue_jays$Mass)
yrng=range(blue_jays$Head)
caption=paste(strwrap("Head length versus body mass for 123 blue jays"))

ggplot(blue_jays, aes(Mass, Head, color=as.factor(Sex)))+
  geom_point(alpha=0.6, size=2)+
  geom_text(data=subset(blue_jays, (Mass>75 & Head>60)|(Head>56 & Head<57 & Mass>62 & Mass<63)), aes(label=KnownSex), nudge_x=0.5)+
  annotate("text", x=xrng[1], y=yrng[2], label=caption, hjust=0, vjust=1, size=4)+
  labs(x="Body mass (g)", y="Head length (mm)")+
  theme(legend.position="none")


```


### Exercise 2

Using `tech_stocks` dataset, recreate the following graphics as precisely as possible.

<br>

##### Plot 1

*Hints:*

- serif font

```{r}

yrng=range(tech_stocks$price_indexed)

xrng=range(tech_stocks$date)

caption=paste(strwrap("Stock price over time for four major tech companies"))




ggplot(tech_stocks, aes(date, price_indexed, color=company))+
  geom_line()+
  geom_text(data=subset(tech_stocks, date==xrng[2]), aes(label=company), family="serif", size=4, color="black", nudge_x = -0.5)+
  annotate("text", x=xrng[1], y=yrng[2], label=caption, hjust=0, vjust=1, size=4, family="serif")+
  labs(x=NULL, y="Stock price, indexed")+
  theme(legend.position="none")



```


<br>

##### Plot 2

*Hints:*

- Package `ggrepel`
- `box.padding` is 0.6
- Minimum segment length is 0 
- Horizontal justification is to the right
- seed of 9876


```{r}

ggplot(tech_stocks, aes(date, price_indexed, color=company))+
  geom_line()+
  geom_text_repel(aes(label=company), data=subset(tech_stocks, date==xrng[2]), hjust=1, box.padding=0.6, color="black", family="serif", segment.colour = "black", min.segment.length=0, show.legend=NA, seed=9876)+
  annotate("text", x=xrng[1], y=yrng[2], label=caption, hjust=0, vjust=1, size=4, family="serif")+
  labs(x=NULL, y="Stock price, indexed")+
  theme(legend.position="none")

```


<br>

### Exercise 3

Using `corruption.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Transparency is 0.6
- Only use 2015 data
- Color used is `grey60`
- `"y ~ log(x)"` (method `"lm"`)
- Package `ggrepel`
- `box.padding` is 0.6
- Minimum segment length is 0 
- seed is 9876

```{r}

corr_sub=corruption %>%
  filter(year==2015) %>%
  filter((cpi>50 & country %in% c("Chile", "Japan", "United States", "Singapore")) | (cpi<50 & country %in% c("Ghana", "Argentina", "Iraq", "Niger", "China")))

xrng=range(corruption$cpi)
yrng=range(corruption$hdi)
  

ggplot(subset(corruption, year==2015), aes(cpi, hdi,color=region))+
  geom_point(alpha=0.6)+
  geom_smooth(aes(group=1), method="lm", formula=y~log(x), se=FALSE, color="grey60")+
  geom_text_repel(aes(label=country), data=corr_sub, box.padding = 0.6, color="black", family="serif", segment.colour = "black", min.segment.length=0, show.legend=NA, seed=9876)+
  annotate("text", x=xrng[1], y=yrng[2], label=caption, hjust=0, vjust=1, size=4, family="serif")+
  labs(x="Corruption Perceptions Index, 2015 (100=least corrupt)", y="Human Development Index, 2015 (1.0=most developed)")


```


<br>

### Exercise 4

Using `cdc` dataset, recreate the described graphic as precisely as possible.

Using Bilbo Baggins' responses below to the CDC BRSFF questions, add Bilbo's data point as a transparent (0.5) solid red circle of size 4 to a scatterplot of `weight` by `height` with transparent (0.1) solid blue circles of size 2 as the plotting characters. In addition, label the point with his name in red. Left justify and rotate the label so it reads vertically from bottom to top --- shift it up by 10 pounds too. Plot should use appropriately formatted axis labels. Remember that the default shape is a solid circle.

* `genhlth` - How would you rate your general health? **fair**
* `exerany` - Have you exercised in the past month? **1=yes**
* `hlthplan` - Do you have some form of health coverage? **0=no**
* `smoke100` - Have you smoked at least 100 cigarettes in your life time? **1=yes** 
* `height` -  height in inches: **46**
* `weight` - weight in pounds: **120**
* `wtdesire` - weight desired in pounds: **120**
* `age` - in years: **45**
* `gender` - m for males and f for females: **m**


Add data
```{r}

bilbo=data.frame("fair", 1, 0,1,46,120,120,45,"m")
names(bilbo)=c("genhlth", "exerany", "hlthplan", "smoke100", "height", "weight", "wtdesire", "age", "gender")

cdc_new=rbind(cdc, bilbo)

```


Plot
```{r}
ggplot(cdc_new, aes(height, weight))+
  geom_point(color="blue", alpha=0.1, size=2)+
  geom_point(data=bilbo, color="red", size=4, alpha=0.5)+
  annotate("text", x=46, y=120+10, label="Bilbo Baggins", hjust=0, size=4, color="red", angle=90)+
  labs(x="Height (inches)", y="Weight (lbs)")


```

<br>







*Hint:* Create a new dataset (maybe call it `bilbo` or `bilbo_baggins`) using either `data.frame()` (base R - example in book) or `tibble()` (tidyverse - see help documentation for the function). Make sure to use variable names that exactly match `cdc`'s variable names.



<br>


# Class Notes #

1. Each layer in ggplot can has its own data

2. We can provide the text layer its own data
