---
title: "L07 Scales, Axes & Legends"
subtitle: "Data Visualization (STAT 302)"
author: "Jingyang Zhang"
output:
  html_document:
    toc: true
    toc_float: true
---


```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE, dpi = 300)
```

## Overview

The goal of this lab is to explore ways to manage and manipulate scales, axes, and legends within `ggplot2`. 

## Datasets

We'll be using the `tech_stocks.rda`, `cdc.txt`, and a few toy datasets.

<br>

```{r load-packages-data-seed, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(lubridate)
# Load datasets
load("data/tech_stocks.rda")

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor")
  ))

# Set seed
set.seed(8221984)

# Selecting a random subset of size 100
cdc_small <- cdc %>% sample_n(100)

# Generating toy datasets for exercise 2
dat1 <- tibble(theta = c(0, 2 * pi))

dat2 <- tibble(
  theta = seq(0, 2 * pi, length.out = 100),
  obs = rnorm(100, sin(theta), 0.1),
  larger_than = ifelse(abs(obs) < abs(sin(theta)), "1", "0")
)
```

<br>

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using the `tech_stocks` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Size of lines in legend is 1.3
- 0.75 and 0.85 will be useful

<br>



Method 1
```{r}
ggplot(tech_stocks, aes(date, price_indexed)) +
  geom_line(aes(color = company)) + 
  scale_color_discrete(limits = c("Facebook", "Alphabet", "Microsoft", "Apple")) + 
  scale_x_date(expand = c(0,0), name = NULL) + 
  scale_y_continuous(breaks = seq(0, 500, 100), position = "right", labels = scales::dollar_format(prefix = "$"), name = NULL) + 
  # Want to override the size of the legend keys
  guides(color = guide_legend(override.aes = list(size = 1.3))) + 
  ggtitle("Stock price, indexed") + 
  theme_minimal() + 
  theme(legend.background=element_blank(), legend.position = c(0.75, 0.85), legend.title = element_blank()) 
  

# Remove legend titles:
#1. theme(legend.title = element_blank())
#2. scale_color_discrete(name = NULL)

# Remove the padding of the graph, meanning that the graph starts and ends at where there is data
#1. scale_x_date(expand = c(0, 0))

# Change the order of colors of the group and legend
#1. scale_color_discrete(limits = c("", "","","",...))

```




Method 2: 
```{r, eval = FALSE}

# Reorder the factor levels
tech_stocks$company <- factor(tech_stocks$company, levels = c("Facebook", "Alphabet", "Microsoft", "Apple"))

# Find color
show_col(hue_pal()(4))


ggplot(tech_stocks, aes(date, price_indexed)) +
  geom_line(aes(color = company)) + 
  scale_color_manual(values = c("Alphabet" = "#7CAE00", "Apple" = "#C77CFF", "Facebook" = "#F8766D", "Microsoft" = "#00BFC4")) + 
  scale_x_date(expand = c(0,0)) + 
  scale_y_continuous(breaks = seq(0, 500, 100), position = "right", labels = scales::dollar_format(prefix = "$")) + 
  xlab(NULL) +
  ylab(NULL) + 
  guides(color = guide_legend(override.aes = list(size = 1.3))) + 
  ggtitle("Stock price, indexed") + 
  theme_minimal() + 
  theme(legend.background=element_blank(), legend.position = c(0.75, 0.85), legend.title = element_blank()) 
``` 





<br>

### Exercise 2

Use the toy datasets `dat1` and `dat2` to recreate the following graphic as precisely as possible.

*Hints:*

- Sizes used 2 (points) and 1.3 (line)
- Transparency 0.8
- Colors: `#56B4E9`, `darkgreen`, & `red`

<br>

```{r}

ggplot(dat2, aes(theta, obs)) + 
  geom_point(aes(color = larger_than), alpha = 0.8, size = 2, show.legend = FALSE) + 
  xlim(dat1$theta) + 
  stat_function(fun = sin, color = "#56B4E9", size = 1.3) + 
  scale_color_manual(values = c("darkgreen", "red")) + 
  labs(x = quote(theta), y = quote(sin(theta))) + 
  theme_minimal()
  
```




<br>

### Exercise 3

Using `cdc_small`construct a scatterplot of `weight` by `height` with the following requirements: 

- Size of plotting characters should be 3.
- Color and shape should both identify `genhlth`.
- One legend for both color and shape.
- Legend title should be "General Health?" with a newline starting after general.
- Legend categories should be ordered from excellent (top) to poor (bottom) with each word in category capitalized in the legend.
- Legend should be placed in the lower right-hand corner of the plotting area. 
- Color should follow the `"Set1"` pallete.
- Shape should have a solid triangle (17) for excellent, solid circle (19) for very good, an `x` (4) for poor, an hollow rotated square with an `x` in it (9) for fair, and a solid square (15) for good.
- `height` values should be limited between 55 and 80.
- `height` axis should display every 5th number between 55 and 80 and be appropriately labeled (i.e. `55 in`, `60 in`, ..., `80 in`). No axis title is necessary.
- `weight` values should be limited between 100 and 300. 
- `weight` axis should be on log base 10 scale, but still display weights in pounds starting at 100 and displaying every 25 pounds until 300. Must be appropriately labeled (i.e. `100 lbs`, `125 lbs`, ..., `300 lbs`). No axis title is necessary.
- Graph title should be `CDC BRFSS: Weight by Height`.
- Minimal theme.

<br>


```{r}

ggplot(cdc_small, aes(height, weight)) + 
  geom_point(aes(color = genhlth, shape = genhlth), 
             size = 3)  + 
  scale_color_brewer(palette = "Set1", labels = c("Excellent", "Very good", "Good", "Fair", "Poor")) +
  labs(color = "General\nHealth?", shape = "General\nHealth?", x = NULL, y = NULL, title = "CDC BRFSS: Weight by Height") + 
  scale_x_continuous(limits = c(55, 80), breaks = seq(55, 80, 5), labels = scales::unit_format(accuracy = 1, unit = "in")) + 
  scale_y_log10(limits = c(100, 300), breaks = seq(100, 300, 25), labels = scales::unit_format(unit = "lbs"))  + 
  scale_shape_manual(labels = c("Excellent", "Very good", "Good", "Fair", "Poor"), values = c(17, 19, 15, 9, 4)) +
  theme_minimal() + 
  theme(legend.position = c(1, 0), legend.justification = c(1, 0))
 


```



## Class Notes ##

1. Each scale_x_x function is built the same
