---
title: "L08 Positioning"
subtitle: "Data Visualization (STAT 302)"
author: "Jingyang Zhang"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, comment = FALSE, dpi = 300)
```

## Overview

The goal of this lab is to develop an understanding of facets, position, continue the exploration of other `ggplot2` options/features. 

## Datasets

We'll be using the `titanic.rda`, `Aus_athletes.rda`, and `cdc.txt` datasets.

<br>

```{r load-packages-data-seed, warning=FALSE, message=FALSE}
# Load package(s)
library(tidyverse)
library(cowplot)

# Load datasets
load("data/titanic.rda")
load("data/Aus_athletes.rda")

# Read in the cdc dataset
cdc <- read_delim(file = "data/cdc.txt", delim = "|") %>%
  mutate(genhlth = factor(genhlth,
    levels = c("excellent", "very good", "good", "fair", "poor"),
    labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")
  ))

# Set seed
set.seed(8221984)

# Selecting a random subset of size 1000
cdc_small <- cdc %>% sample_n(1000)
```

<br>

## Exercises

Complete the following exercises.

<br>

### Exercise 1

Using the `titanic.rda` dataset, recreate the following graphic as precisely as possible.

*Hints:*

- Hex colors: `#D55E00D0`, `#0072B2D0`

<br>

```{r}

# New facet label names for dose variable
survival.labs <- c("died", "survived")
names(survival.labs) <- c("0", "1")

ggplot(titanic, aes(sex)) + 
  geom_bar(aes(fill = sex), show.legend = FALSE) + 
  facet_grid(survived ~ class, labeller = labeller(survived = survival.labs)) +
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) + 
  theme_minimal() 
  

  
  

  

```




<br>

### Exercise 2

Use the `athletes_dat` dataset --- extracted from `Aus_althetes.rd` --- to recreate the following graphic as precisely as possible. The `cowplot` package will be useful.

<br>

```{r}
# Get list of sports played by BOTH sexes
both_sports <- Aus_athletes %>%
  distinct(sex, sport) %>%
  count(sport) %>%
  filter(n == 2) %>%
  pull(sport)

# Process data
athletes_dat <- Aus_athletes %>%
  filter(sport %in% both_sports) %>%
  mutate(sport = case_when(
    sport == "track (400m)" ~ "track",
    sport == "track (sprint)" ~ "track",
    TRUE ~ sport
    )
  )
```

<br>

*Hints:*

- Build each plot separately
- Use [`cowplot::plot_grid()`](https://cran.r-project.org/web/packages/cowplot/vignettes/plot_grid.html) to combine them
- Hex values for shading: `#D55E0040` and `#0072B240` (bottom plot),
`#D55E00D0` & `#0072B2D0` (for top two plots) --- no `alpha`
- Hex values for outline of boxplots: `#D55E00` and `#0072B2`
- Boxplots should be made narrower; 0.5
- Legend is in top-right corner of bottom plot
- Legend shading matches hex values for top two plots
- Bar plot lower limit 0, upper limit 95
- `rcc`: red blood cell count; `wcc`: white blood cell count
- Size 3 will be useful

<br>

```{r}


exer2_p1 <- ggplot(athletes_dat, aes(sex)) +
  geom_bar(aes(fill = sex), show.legend = FALSE) + 
  scale_x_discrete(labels = c("female", "male"), name = NULL) +
  scale_y_continuous(limits = c(0, 95), breaks = seq(0, 95, 25), name = "number") + 
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  theme_minimal()

exer2_p2 <- ggplot(athletes_dat, aes(rcc, wcc)) +  
  geom_point(aes(color = sex, fill = sex), shape = 21,
             size = 3, color = "white", show.legend = FALSE)+
  scale_fill_manual(values = c("#D55E00D0", "#0072B2D0")) +
  labs(x = "RBC count", y = "WBC count") + 
  theme_minimal()


exer2_p3 <- ggplot(athletes_dat, aes(sport, pcBfat)) + 
  geom_boxplot(aes(fill = sex, 
                   color = sex), width = 0.5) +
  scale_color_manual(values = c("#D55E00", "#0072B2"), 
                     labels = c("female", "male"), 
                     # guide = FALSE is NOT necessary,
                     # it is used to turn off one of the legend
                     guide = FALSE ) +
  # the 40 at the end of #D55E00 adds the transparency
  scale_fill_manual(values = c("#D55E0040", "#0072B240"),   
                    labels = c("female", "male")) + 
  guides(color = guide_legend(override.aes = list(
    color = NA, # remove the outline of boxplot
    fill = c("#D55E00D0", "#0072B2D0")
    ))) +
  labs(fill = NULL,  color = NULL, 
       y = "% body fat", x = NULL) +
theme_minimal() + 
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1), 
        legend.direction = "horizontal")

  

exer2_row1 <- plot_grid(exer2_p1, exer2_p2)

plot_grid(exer2_row1, exer2_p3, ncol = 1)

```


Method 2 of making exer2_p3
```{r, eval = FALSE}
exer2_p3 <- ggplot(athletes_dat, aes(sport, pcBfat)) + 
  geom_boxplot(aes(fill = sex, 
                   color = sex), width = 0.5) +
  scale_color_manual(values = c("#D55E00", "#0072B2"), 
                     labels = c("female", "male")) +
  # the 40 at the end of #D55E00 adds the transparency
  scale_fill_manual(values = c("#D55E0040", "#0072B240"),   
                    labels = c("female", "male")) + 
  guides(fill = guide_legend(override.aes = list(fill = c("#D55E00D0", "#0072B2D0"))), color = guide_legend(override.aes = list(color = c("transparent", "transparent")))) +
  labs(fill = NULL,  color = NULL, 
       y = "% body fat", x = NULL) +
theme_minimal() + 
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1), 
        legend.direction = "horizontal")
```



<br>

### Exercise 3

Using `cdc_small`construct a the following graphic as precisely as possible.

*Hints:*

- Need to do a little more data processing
- Hex values: `"#D55E00D0"` and `#0072B2D0`
- `grey80` might be useful
- size 2

<br>

```{r}
# Add calculated weight change
# Reorder genhlth levels
cdc_small <- cdc_small %>%
  mutate(
    weight_change = wtdesire - weight,
    genhlth_order= factor(genhlth, levels = rev(levels(factor(genhlth))))
  )

gender.labs <- c("Women", "Men")
names(gender.labs) <- c("f", "m")

cdc_small2 <- select(cdc_small, -gender)

ggplot(cdc_small, aes(weight, weight_change)) +
  geom_point(data = cdc_small2, color = "grey80") +
  geom_point(aes(color = gender), size = 2, show.legend = FALSE) +
  facet_grid(gender ~ genhlth_order, labeller = labeller(gender = gender.labs)) +
  scale_color_manual(values = c("#D55E00D0", "#0072B2D0")) + 
  labs(x = "Weight (lbs)" , y = "Weight Loss/Gain in Pounds") +
  theme_minimal()





```




<br>
